---
title: "HW3"
author: "Roxana Chicas"
date: "February 8, 2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
wine <- read.csv("White_wines.csv",
               header = TRUE)

```

```{r}
# see what is in the White_wine dataset
summary(wine)
```

```{r}
# divide the dataset into a training and a testing set based on a random uniform number on fixed seed
set.seed(20170208)
wine$group <- runif(length(wine$quality), min = 0, max = 1)
summary(wine)
wine.train <- subset(wine, group <= 0.90)
wine.test <- subset(wine, group > 0.90)

```

Now let's graph our training set data.
```{r}
#graph fixed.acidity on quality for testing dataset
plot(quality ~ fixed.acidity, data = wine.train)
```
```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(fixed.acidity)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$fixed.acidity), 
                              sd = sd(wine.train$fixed.acidity)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of Fixed Acidity",
       x = "fixed.acidity",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(volatile.acidity)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$volatile.acidity), 
                              sd = sd(wine.train$volatile.acidity)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of Volatile Acidity",
       x = "volatile.acidity",
       y = "Density")
```
```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(citric.acid)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$citric.acid), 
                              sd = sd(wine.train$citric.acid)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of citric acid",
       x = "citric.acid",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(residual.sugar)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$residual.sugar), 
                              sd = sd(wine.train$residual.sugar)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of residual sugar",
       x = "residual.sugar",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(chlorides)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$chlorides), 
                              sd = sd(wine.train$chlorides)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of chlorides",
       x = "chlorides",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(free.sulfur.dioxide)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$free.sulfur.dioxide), 
                              sd = sd(wine.train$free.sulfur.dioxide)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of free sulfur dioxide",
       x = "free.sulfur.dioxide",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(total.sulfur.dioxide)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$total.sulfur.dioxide), 
                              sd = sd(wine.train$total.sulfur.dioxide)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of total sulfur dioxide",
       x = "total.sulfur.dioxide",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(total.sulfur.dioxide)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$total.sulfur.dioxide), 
                              sd = sd(wine.train$total.sulfur.dioxide)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of total sulfur dioxide",
       x = "total.sulfur.dioxide",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(density)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$density), 
                              sd = sd(wine.train$density)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of density",
       x = "density",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(pH)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$pH), 
                              sd = sd(wine.train$pH)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distrubution of pH",
       x = "pH",
       y = "Density")
```
```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(sulphates)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$sulphates), 
                              sd = sd(wine.train$sulphates)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of sulphates",
       x = "Sulfphates",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(alcohol)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$alcohol), 
                              sd = sd(wine.train$alcohol)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of alcohol",
       x = "alcohol",
       y = "Density")
```

```{r}
library(magrittr)
library(ggplot2)
wine.train %>%
  ggplot(aes(quality)) +
    geom_histogram(aes(y=..density..),
                   colour="black",fill="yellow",
                   binwidth=5) + 
    stat_function(fun = dnorm, 
                  args = list(mean = mean(wine.train$quality), 
                              sd = sd(wine.train$quality)), 
                  lwd = 1, 
                  col = 'red') +
  labs(title = "Distribution of quality",
       x = "quality",
       y = "Density")
```

```{r}
wine.train %>%
  ggplot(aes(x=fixed.acidity, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ fixed.acidity, data = wine.train)
  summary (reg2)
```


```{r}
wine.train %>%
  ggplot(aes(x=volatile.acidity, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ volatile.acidity, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=citric.acid, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ citric.acid, data = wine.train)
  summary (reg2)
```
```{r}
wine.train %>%
  ggplot(aes(x=residual.sugar, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ residual.sugar, data = wine.train)
  summary (reg2)
```
```{r}
wine.train %>%
  ggplot(aes(x=chlorides, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ chlorides, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=free.sulfur.dioxide, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ free.sulfur.dioxide, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=total.sulfur.dioxide, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ total.sulfur.dioxide, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=density, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ density, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=pH, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ pH, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=sulphates, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ sulphates, data = wine.train)
  summary (reg2)
```

```{r}
wine.train %>%
  ggplot(aes(x=alcohol, y=quality)) +
    geom_point(position = "jitter")
```
```{r}
  reg2 <-lm(quality ~ alcohol, data = wine.train)
  summary (reg2)
```

```{r}
library(car)
#produce a scatterplot matrix 
scatterplotMatrix(~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, span =0.5, data = wine)
```
the variables "residual sugar" and "chlorides" look kind of wonky so transformation may help.

```{r}
library(car)
scatterplotMatrix(~ fixed.acidity + volatile.acidity + citric.acid + log2(residual.sugar) + log2(chlorides) + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, span =0.5, data = wine)
```


```{r}
# run regression with all variables and transformation
wine.mod1 <- lm(quality ~  fixed.acidity + volatile.acidity + citric.acid + log2(residual.sugar) + log2(chlorides) + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data= wine.train)

#and see what we have
summary(wine.mod1)
```

```{r}
# regression ripped without non-significant variables
wine.mod2 <- lm(quality ~ volatile.acidity + log2(residual.sugar) + log2(chlorides) + free.sulfur.dioxide + density + pH + sulphates + alcohol, data= wine.train)

#and here we have it
summary(wine.mod2)
```
The ajusted R-square remained the same in both models 0.2777.  The second model therefore is more parsomonious since it includes less variables.  

```{r results="asis"}
library(stargazer)
# compare the results of the two regression models
stargazer(wine.mod1,wine.mod2,title="Comparison of 2 Regression outputs",type="html",align=TRUE)
```

```{r}
# diagnostics for the first model
residualPlots(wine.mod1)

```
```{r}
#added variable plots
avPlots(wine.mod1, id.n=2, id.cex=0.7)
#id.n - identify n most influential observations
#id.cex - controls the size of the dot
```
```{r}
vif(wine.mod1)
```
```{r}
#make influence plot
influencePlot(wine.mod1, id.n=3)
```
```{r}

# diagnostics for the second model 
residualPlots(wine.mod2)
```
```{r}
#added variable plots
avPlots(wine.mod2, id.n=2, id.cex=0.7)
#id.n - identify n most influential observations
#id.cex - controls the size of the dot
```
```{r}
vif(wine.mod2)
```
```{r}
#make influence plot
influencePlot(wine.mod2, id.n=3)
```
```{r}
# regression ripped without non-significant variables and removed variables with vif >4
wine.mod3 <- lm(quality ~ volatile.acidity + log2(chlorides) + free.sulfur.dioxide + pH + sulphates, data= wine.train)

#and here we have it
summary(wine.mod3)
```
Once vif >4 removed from model 2 "free sulfur dioxide" became non-significant and ajusted R-square reduced to 0.1156 from 0.2777 in model 2.


Based on the the three models ran, the second model is the better model.  Model 2 does not have non-significant variables and the ajusted R-square is 0.2777.  Model 1 and 2 have the same R-square but model 2 is parsimonious since it has feweer variables.  Model three was not used because the ajusted R-square decrease dramtically once vif >4 were removed.  According to the output in model 2 for every one unit decrease of log2(residual.sugar) there is an increase in wine quality rating  of 1.927e-01 when all other variables are held constant.  Output in model 2 also shows that for every one unit decrease of log2(chlorides) there is an decrease in wine quality rating  of -6.224e-02 when all other variables are held constant.  A one unit increase of free sulfurdioxide increases wine qulity rating of 3.003e-03 when all variables are held constant.  As one unit of density increases there is wine quality decreases of -7.381e+01.  pH showed that for every one unit increase the wine quality increased by 3.691e-01.  Sulphates show that for one unit increase there is an increase in wine quality of 4.995e-01.  Model 2 output showed that for alcohol increase of 1 unit there is an increase of wine quality of 2.584e-01.   The multiple R-squared is 0.272, indicating that 27% of the variability in wine quality rating is due to these eight independent variables.

The End :)